Topological Equivariant Artist Model for Visualization Library Architecture

### 2024-11-26 paper/main.tex
homeomorphism figure

As illustrated in \ref{fig:def:homomorphism}, .... structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2. We must define a < for visual compatable with the < for numbers. We handwave that for every relation we can write a compatabile closed function,  e.g. we can replace $\leq$ with $\min$, such that the operation is closed.

@@

\section{Codifying Structure Preservation}
\label{sec:artist}
In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
\end{align}
 The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
 \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
\begin{itemize}
  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$
\end{itemize}
such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
 \end{definition}
 This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
  \label{eq:artist:sheaf:base}
    \begin{tikzcd}
      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
      \end{tikzcd}
\end{equation}
take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and
\autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:

>>>>

As illustrated in \ref{fig:def:homomorphism}, structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2. We must define a < for visual compatable with the < for numbers. We handwave that for every relation we can write a compatabile closed function,  e.g. we can replace $\leq$ with $\min$, such that the operation is closed.

@@

\section{Codifying Structure Preservation}
\label{sec:artist}
In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
\end{align}
 The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
 \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
\begin{itemize}
  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$
\end{itemize}
such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
 \end{definition}
 This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
  \label{eq:artist:sheaf:base}
    \begin{tikzcd}
      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
      \end{tikzcd}
\end{equation}
take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and
\autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
\note{this needs human words - subsets of functions of the same type map to subsets of visualizations of the same type}
